---
title: "Threatened-species-bycatch-hotspots-in-the-tropical-Atlantic-Ocean"
author: "Clara Lerebourg"
format: html
editor: visual
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
#| label: setup
#| include: false

# Load the necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(vegan)
library(RPostgres)
library(knitrdata)
library(sf, quietly=TRUE)
library(units)
library(patchwork)
library(igraph)

# parameters
start_date = '2013-01-01'
end_date = '2022-12-31'
world_boundaries <- rnaturalearth::ne_countries(returnclass = "sf",
                                                scale       = "medium")
ocean_xlim <- c(-40, 15)
ocean_ylim <- c(-25, 25)
ocean_xlim_short <- c(-13, 15)
ocean_ylim_short <- c(-20, 8)
ocean_xintercept <- c(-30, -20, -10, 0, 10)
ocean_yintercept <- c(-20, -10, 0, 10, 20)
```

# *All the functions*

## cwp_type_to_scales

```{r}
cwp_type_to_scales <- function(type) {
  type <- as.character(type)

  ll <- data.frame(
    "5" = c(1, 1), "11" = c(1, 1),
    "6" = c(5, 5), "55" = c(5, 5),
    "1" = c(10, 5), "105" = c(10, 5),
    "7" = c(10, 10), "1010" = c(10, 10),
    "2" = c(20, 10), "2010" = c(20, 10),
    "8" = c(20, 20), "2020" = c(20, 20),
    "9" = c(30, 30), "3030" = c(30, 30),
    check.names = FALSE
  )
  ll <- as.data.frame(t(ll))
  names(ll) <- c("lonscale", "latscale")

  I <- type %in% row.names(ll)
  if (!all(I)) {
    stop("Unknown CWP type: ", type[!I][1])
  }

  return(ll[type, , drop = FALSE])
}
```

## cwp_code_to_geom

```{r, message=FALSE, warning=FALSE, include=FALSE}
cwp_code_to_geom <- function(cwp,
                             type = NULL,
                             centroid = TRUE) {
  cwp <- as.character(cwp)
  if (is.null(type)) {
    if (!all(nchar(cwp) == 7)) {
      stop("Poorly formed cwp or type. Possibly cwp missing type.")
    }
    type <- as.numeric(substr(cwp, 1, 1))
    cwp <- substr(cwp, 2, 1000)
  } else {
    if (!all(nchar(cwp) == 6)) {
      stop("Poorly formed cwp or type. Possibly cwp including type.")
    }
  }
  scales <- cwp_type_to_scales(type)
  quad <- as.integer(substr(cwp, 1, 1))
  signs <- rbind(c(1, 1), c(1, -1), c(-1, -1), c(-1, 1))
  signs <- signs[quad, , drop = FALSE]
  lat0 <- as.integer(substr(cwp, 2, 3))
  lon0 <- as.integer(substr(cwp, 4, 6))
  lon <- signs[, 1] * lon0
  lat <- signs[, 2] * lat0
  if (centroid) {
    ln <- signs[, 1] * (lon0 + 0.5 * scales[, 1])
    lt <- signs[, 2] * (lat0 + 0.5 * scales[, 2])
    # This seems like a hard way to make points
    pt <- sapply(
      seq_along(ln),
      function(i) {
        sf::st_point(c(ln[i], lt[i]))
      }, simplify = FALSE
    )
    pt <- sf::st_sfc(pt, crs = 4326)
  } else {
    ln <- signs[, 1] * (lon0 + scales[, 1])
    lt <- signs[, 2] * (lat0 + scales[, 2])

    # This seems like a hard way to make polygons
    pt <- sapply(
      seq_along(ln),
      function(i) {
        x <- cbind(
          c(lon[i], ln[i], ln[i], lon[i], lon[i]),
          c(lat[i], lat[i], lt[i], lt[i], lat[i])
        )
        sf::st_polygon(list(x))
      }, simplify = FALSE
    )
    pt <- sf::st_sfc(pt, crs = 4326)
  }
  return(pt)
}
```

## pie_chart_species_composition

```{r, message=FALSE, warning=FALSE, include=FALSE}
pie_chart_species_composition <- function(dataframe,
                                          threshold = 9) {
  # Filtering data for the hotspot
  # Selection of species columns
  species_counts <- dataframe[, 3:12] %>%
    dplyr::summarise(dplyr::across(tidyr::everything(),
                                   sum))
  
  # Conversion to long format
  species_df <- data.frame(species = names(species_counts),
                           count = as.numeric(species_counts)) %>%
    dplyr::mutate(percentage = (count / sum(count)) * 100) %>%
    dplyr::mutate(species = ifelse(percentage < threshold, 
                                   "Other",
                                   as.character(species))) %>%
    dplyr::group_by(species) %>%
    dplyr::summarise(count = sum(count),
                     .groups = "drop") %>%
    dplyr::mutate(percentage = (count / sum(count)) * 100) %>%
    dplyr:: mutate(species = factor(species,
                                    levels = c(setdiff(species[species != "Other"], 
                                                       "Other"), "Other"))) %>%
    dplyr::arrange(percentage)
  
  # Graphic design
  pie_chart_plot <- ggplot2::ggplot(species_df, ggplot2::aes(x = "",
                                                             y = percentage,
                                                             fill = species)) +
    ggplot2::geom_bar(stat = "identity",
                      width = 1) +
    ggplot2::coord_polar(theta = "y") +
    ggplot2::scale_fill_manual(values = c("Billfishes" = "#F07167",
                                          "Devil_ray" = "darkorchid",
                                          "Blue_shark" = "darkturquoise",
                                          "Silky_shark" = "#00AFB9",
                                          "Shortfin_Mako" = "blue4",
                                          "Olive_ridley_turtle" = "#ffac81",
                                          "Loggerhead_turtle" = "#FED9B7",
                                          "Scalloped_hammerhead" = "#0081A7",
                                          "Smooth_hammerhead" = "cadetblue2",
                                          "Other" = "grey")) +
    ggplot2::labs(title = "", 
                  fill = "species") +
    ggplot2::theme_void()
  return(pie_chart_plot)
}
```

## summarize_data

```{r, message=FALSE, warning=FALSE, include=FALSE}
summarize_data <- function(dataframe,
                           size_juv) {
  # Calculation of proportions by category
  nb_male_juvenile <- round(sum(dataframe$sex == "Male" &
                                  dataframe$length <= size_juv,
                                na.rm = TRUE))
  nb_female_juvenile <- round(sum(dataframe$sex == "Female"
                                  & dataframe$length <= size_juv,
                                  na.rm = TRUE))
  nb_male_adult <- round(sum(dataframe$sex == "Male"
                             & dataframe$length > size_juv,
                             na.rm = TRUE))
  nb_female_adult <- round(sum(dataframe$sex == "Female"
                               & dataframe$length > size_juv,
                               na.rm = TRUE))
  return(data.frame(nb_male_juvenile,
                    nb_female_juvenile,
                    nb_male_adult,
                    nb_female_adult))
}
```

## summarize_percentage

```{r, message=FALSE, warning=FALSE, include=FALSE}
summarize_percentage <- function(data,
                                 size_juv) {
  data_filtered <- data %>% dplyr::filter(sex %in% c("Male",
                                                     "Female"))
  total_ind <- sum(data_filtered$count, 
                   na.rm = TRUE)
  # Calculation of percentage by category
  perc_male_juvenile <- round(sum(data$sex == "Male" & 
                                    data$length <= size_juv, 
                                  na.rm = TRUE) / total_ind * 100)
  perc_female_juvenile <- round(sum(data$sex == "Female" & 
                                      data$length <= size_juv, 
                                    na.rm = TRUE) / total_ind * 100)
  perc_male_adult <- round(sum(data$sex == "Male" & 
                                 data$length > size_juv, 
                               na.rm = TRUE) / total_ind * 100)
  perc_female_adult <- round(sum(data$sex == "Female" & 
                                   data$length > size_juv, 
                                 na.rm = TRUE) / total_ind * 100)
  return(data.frame(perc_male_juvenile,
                    perc_female_juvenile,
                    perc_male_adult,
                    perc_female_adult,
                    total_ind))
}
```

## pie_chart_percentage_bio

```{r, message=FALSE, warning=FALSE, include=FALSE}
pie_chart_percentage_bio <- function(dataframe,
                                     title,
                                     legend = FALSE) {
  if ("total_ind" %in% names(dataframe)) {
    dataframe <- dataframe %>% dplyr::select(-total_ind)
    }
  data_long <- dataframe %>%
    tidyr::pivot_longer(cols = tidyr::everything(),
                        names_to = "category",
                        values_to = "percentage")
  if (legend == TRUE) {
    ggplot2::ggplot(data_long %>%

                      ggplot2::aes(x = "",
                                   y = percentage,
                                   fill = category)) +
      ggplot2::geom_bar(stat = "identity", width = 1) +
      ggplot2::coord_polar("y", start = 0) +
      ggplot2::theme_minimal() +
      ggplot2::theme(axis.text.x = ggplot2::element_blank(),
                     axis.ticks = ggplot2::element_blank(),
                     panel.grid = ggplot2::element_blank()) +
      ggplot2::labs(title = title, y = NULL, x = NULL) +
      ggplot2::scale_fill_manual(values = c(perc_male_adult = "#005f73",
                                            perc_male_juvenile = "#0a9396",
                                            perc_female_adult = "#B56B3C",
                                            perc_female_juvenile = "#e9d8a6"))
  } else {
    ggplot2::ggplot(data_long,
                    ggplot2::aes(x = "",
                                 y = percentage,
                                 fill = category)) +
      ggplot2::geom_bar(stat = "identity", width = 1) +
      ggplot2::coord_polar("y", start = 0) +
      ggplot2::theme_minimal() +
      ggplot2::theme(axis.text.x = ggplot2::element_blank(),
                     axis.ticks = ggplot2::element_blank(),
                     panel.grid = ggplot2::element_blank(),
                     legend.position = "none") +
      ggplot2::labs(title = title, y = NULL, x = NULL) +
      ggplot2::scale_fill_manual(values = c(perc_male_adult = "#005f73",
                                            perc_male_juvenile = "#0a9396",
                                            perc_female_adult = "#B56B3C",
                                            perc_female_juvenile = "#e9d8a6"))
  }
}
```

## generate_fishing_maps_and_charts

```{r, message=FALSE, warning=FALSE, include=FALSE}
generate_fishing_maps_and_charts <- function(dataframe,
                                             world_boundaries,
                                             hotspot,
                                             species) {
  # Select column by species
  species_col <- paste0("num_", species)
  # A - Preparing the data for the map
  bs <- dataframe %>%
    dplyr::select(the_geom,
                  annee_de_peche,
                  mois_de_peche,
                  tidyr::all_of(species_col)) %>%
    dplyr::filter(.data[[species_col]] >= 1) %>%
    dplyr::mutate(longitude = sf::st_coordinates(the_geom)[, 1],
                  latitude = sf::st_coordinates(the_geom)[, 2]) %>%
    sf::st_drop_geometry()
  
  bs$annee_de_peche <- factor(bs$annee_de_peche,
                              levels = 2013:2022,
                              ordered = TRUE)
  bs$mois_de_peche <- factor(lubridate::month(bs$mois_de_peche),
                             levels = 1:12,
                             ordered = TRUE)
  # Definition of geographical limits
  hotspot_limits <- list(mauritania = list(xlim = c(-25, -10),
                                           ylim = c(10, 22)),
                         gabon = list(xlim = c(0, 15),
                                      ylim = c(5, -8)),
                         angola = list(xlim = c(20, 5),
                                       ylim = c(-5, -17)))[[hotspot]]
  
  ocean_xintercept <- c(-30, -20, -10, 0, 10)
  ocean_yintercept <- c(-20, -10, 0, 10, 20)
  
  # B - Preparation of data for the total number of sets
  bs_total_sets <- dataframe %>%
    dplyr::select(the_geom,
                  obs_set_id,
                  annee_de_peche,
                  mois_de_peche,
                  nombre_de_calees_pos) %>%
    dplyr::mutate(longitude = sf::st_coordinates(the_geom)[, 1],
                  latitude = sf::st_coordinates(the_geom)[, 2]) %>%
    sf::st_drop_geometry()
  # Addition of the nombre_de_calees column (number of sets)
  data_set_group <- bs_total_sets %>%
    dplyr::group_by(obs_set_id,
                    annee_de_peche,
                    mois_de_peche) %>%
    dplyr::summarize(nombre_de_calees = sum(nombre_de_calees_pos,
                                            na.rm = TRUE),
                     longitude = dplyr::first(longitude),
                     latitude = dplyr::first(latitude))
  # Factorisation
  data_set_group$annee_de_peche <- factor(data_set_group$annee_de_peche,
                                          levels = 2013:2022,
                                          ordered = TRUE)
  data_set_group$mois_de_peche <- lubridate::month(data_set_group$mois_de_peche)
  data_set_group$mois_de_peche <- factor(data_set_group$mois_de_peche,
                                         levels = 1:12,
                                         ordered = TRUE)
  
  # Preparation of data for the number of sets with the species
  bs_species <- bs %>%
    dplyr::filter(dplyr::case_when(hotspot == "mauritania" ~ latitude >= 16 & latitude <= 21 & 
                                     longitude >= -21 & longitude <= -16,
                                   hotspot == "gabon" ~ latitude <= 0 & latitude >= -4 & 
                                     longitude >= 7 & longitude <= 10,
                                   hotspot == "angola" ~ latitude <= -7 & latitude >= -13 & 
                                     longitude >= 10 & longitude <= 14,
                                   TRUE ~ FALSE))
  
  data_set_group <- data_set_group %>%
    dplyr::filter(dplyr::case_when(hotspot == "mauritania" ~ latitude >= 16 & latitude <= 21 & 
                                     longitude >= -21 & longitude <= -16,
                                   hotspot == "gabon" ~ latitude <= 0 & latitude >= -4 & 
                                     longitude >= 7 & longitude <= 10,
                                   hotspot == "angola" ~ latitude <= -7 & latitude >= -13 & 
                                     longitude >= 10 & longitude <= 14,
                                   TRUE ~ FALSE))
  
  # Count of sets with the selected species and total
  # Calculation of the number of sets with species per year
  species_counts_year <- bs_species %>%
    dplyr::group_by(annee_de_peche) %>%
    dplyr::summarize(count_species = dplyr::n())
  
  species_counts_month <- bs_species %>%
    dplyr::group_by(mois_de_peche) %>%
    dplyr::summarize(count_species = dplyr::n())
  
  # Calculation of the total number of sets per year
  set_counts_year <- data_set_group %>%
    dplyr::group_by(annee_de_peche) %>%
    dplyr::summarize(count_total_sets = dplyr::n())
  
  set_counts_month <- data_set_group %>%
    dplyr::group_by(mois_de_peche) %>%
    dplyr::summarize(count_total_sets = dplyr::n())
  
  # YEAR - Combination of the two data sets
  combined_counts_year <- dplyr::full_join(species_counts_year,
                                           set_counts_year,
                                           by = "annee_de_peche")
  combined_counts_year_long <- combined_counts_year %>%
    tidyr::gather(key = "Type",
                  value = "Count",
                  count_species,
                  count_total_sets)
  
  # MONTH - Combination of the two data sets
  combined_counts_month <- dplyr::full_join(species_counts_month,
                                            set_counts_month,
                                            by = "mois_de_peche")
  combined_counts_month_long <- combined_counts_month %>%
    tidyr::gather(key = "Type",
                  value = "Count",
                  count_species,
                  count_total_sets)
  
  # Graphic design 
  # Map of sets with the selected species
  bs <- bs %>%
    dplyr::arrange(.data[[species_col]])
  point_witout_species <- dataframe %>%
    dplyr::select(the_geom,
                  tidyr::all_of(species_col)) %>%
    dplyr::filter(.data[[species_col]] == 0) %>%
    dplyr::mutate(longitude = sf::st_coordinates(the_geom)[, 1],
                  latitude = sf::st_coordinates(the_geom)[, 2]) %>%
    sf::st_drop_geometry()
  (map <- ggplot2::ggplot() +
      ggplot2::theme(legend.text = ggplot2::element_text(size = 10),
                     panel.background = ggplot2::element_rect(fill = "white"),
                     panel.border = ggplot2::element_rect(color = "black", 
                                                          fill = NA,
                                                          size = 0.3),
                     axis.title = ggplot2::element_text(size = 12),
                     axis.text = ggplot2::element_text(size = 10),
                     legend.position = c(0.15, 0.13),
                     legend.box = "horizontal") +  # Légendes côte à côte
      ggplot2::geom_sf(data = world_boundaries) +
      ggspatial::coord_sf(xlim = hotspot_limits$xlim,
                          ylim = hotspot_limits$ylim) +
      ggplot2::geom_point(data = point_witout_species,
                          mapping = ggplot2::aes(x = longitude, 
                                                 y = latitude),
                          color = "grey",
                          size = 1,
                          show.legend = FALSE) +
      ggplot2::geom_point(data = bs,
                          ggplot2::aes(x = longitude,
                                       y = latitude,
                                       size = .data[[species_col]],
                                       color = .data[[species_col]]),
                          show.legend = c(color = TRUE, size = TRUE)) +
      ggplot2::geom_hline(yintercept = ocean_yintercept,
                          linetype = "dashed",
                          color = "darkgrey",
                          linewidth = 0.2) +
      ggplot2::geom_vline(xintercept = ocean_xintercept,
                          linetype = "dashed",
                          color = "darkgrey",
                          linewidth = 0.2) +
      ggplot2::scale_x_continuous(breaks = seq(-40, 10, 
                                               by = 10)) +
      ggplot2::scale_y_continuous(breaks = seq(-20, 20, 
                                               by = 10)) +
      ggplot2::scale_color_viridis_c(option = "plasma") +
      ggplot2::scale_size_continuous(breaks = c(40, 30, 20, 10),  
                                     labels = c("40", "30", "20", "10"),
                                     guide = ggplot2::guide_legend(title.position = "top")) +
      ggplot2::guides(color = ggplot2::guide_colorbar(title.position = "top"),
                      size = ggplot2::guide_legend(title.position = "top")) +
      ggplot2::labs( caption = "",
                     x = "",
                     y = ""))
  
  # Set by year
  ylim_p1 <- max(combined_counts_year_long$Count, na.rm = TRUE) + 10
  (plot1 <- ggplot2::ggplot(combined_counts_year_long,
                            ggplot2::aes(x = annee_de_peche,
                                         y = Count,
                                         fill = Type)) +
      ggplot2::geom_bar(stat = "identity",
                        position = "dodge") +
      ggplot2::geom_text(ggplot2::aes(label = Count),
                         position = ggplot2::position_dodge(width = 0.9),
                         vjust = -0.5,
                         size = 4) +  # Add values above the bars
      ggplot2::labs(x = "Year",
                    y = "Count",
                    fill = "") +  
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = c(0.05, 0.95),  # Legend top left
                     legend.justification = c("left", "top"),
                     legend.text = ggplot2::element_text(size = 20)) +  # Increases the size of the caption title
      ggplot2::scale_x_discrete(drop = FALSE) +
      ggplot2::ylim(0, ylim_p1) +
      ggplot2::scale_fill_manual(values = c("#9B59B6",
                                            "#FFCC00"),  
                                 labels = c("Count of sets with species",
                                            "Count of sets total"),
                                 name = "") +  
      ggplot2::theme(legend.position = "none",
        legend.text = ggplot2::element_text(size = 10),
        axis.text.x = ggplot2::element_text(size = 10),  # Increases the text size of the x axis graduations
        axis.text.y = ggplot2::element_text(size = 10), # Increases the text size of the y axis graduations
        axis.title.x = ggplot2::element_text(size = 10), # Increases the size of the x-axis title
        axis.title.y = ggplot2::element_text(size = 10))  # Increases the size of the y-axis title
  )
  
  # Set by month
  ylim_p2 <- max(combined_counts_month_long$Count, na.rm = TRUE) + 10
  (plot2 <- ggplot2::ggplot(combined_counts_month_long,
                            ggplot2::aes(x = mois_de_peche,
                                         y = Count,
                                         fill = Type)) +
      ggplot2::geom_bar(stat = "identity",
                        position = "dodge") +
      ggplot2::geom_text(ggplot2::aes(label = Count),
                         position = ggplot2::position_dodge(width = 0.9),
                         vjust = -0.5,
                         size = 4) +  
      ggplot2::labs(x = "Month",
                    y = "Count",
                    fill = "") +  #
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = c(0.05, 0.95),  
                     legend.justification = c("left", "top"),
                     legend.text = ggplot2::element_text(size = 15)) +  
      ggplot2:: scale_x_discrete(drop = FALSE) +
      ggplot2::ylim(0, ylim_p2) +
      ggplot2::scale_fill_manual(values = c("#9B59B6",
                                            "#FFCC00"),  
                                 labels = c("Count of sets with species",
                                            "Count of sets total"),
                                 name = "") +  
      ggplot2::theme(#legend.position = "none",
                     axis.text.x = ggplot2::element_text(size = 10),  
                     axis.text.y = ggplot2::element_text(size = 10),  
                     axis.title.x = ggplot2::element_text(size = 10), 
                     axis.title.y = ggplot2::element_text(size = 10))
  )
  
  # Adjusting the theme for a longer map
  map <- map + ggplot2::theme(aspect.ratio = 0.75,
                              legend.text = ggplot2::element_text(size = 12))
  
  # Combination of the two bar graphs (plot1 above plot2)
  plots_right <- cowplot::plot_grid(plot1,
                                    plot2,
                                    ncol = 1,
                                    align = "v",
                                    rel_heights = c(1, 1))
  
  # Organisation finale avec la carte plus large à gauche
  final_plot <- cowplot::plot_grid(map,
                                   plots_right,
                                   ncol = 2,
                                   align = "h",
                                   rel_widths = c(1.2, 1))
  
  # Print
  return(final_plot)
}

```

# *Clean the data*

```{r, message=FALSE, warning=FALSE, include=FALSE}
#| label: load_data
#| echo: true

# ORIGINAL DATA ----
data <- readRDS("E:/git/reduce/data/ecd_wide_obs_fr_catch_sets_data.RDS")
# Addition of columns cwp11_geom and cwp55_geom
data <- data %>%
  dplyr::mutate("year" = lubridate::year(x = d_act),
                "cwp11_geom" = cwp_code_to_geom(data$cwp11,
                                                        type = 11,
                                                        centroid = FALSE),
                "cwp55_geom" = cwp_code_to_geom(data$cwp55,
                                                        type = 55,
                                                        centroid = FALSE))
data <- as.data.frame(data)
# Filter the data : Atlantic, sets with observer data, ..
obs_clean <- data %>%
  dplyr::filter(ocean == "Atlantic", 
                !is.na(obs_set_id), # Only sets with observer data
                nombre_de_calees_pos > 0, # Only non-null sets
                obs_school_type_code == code_assoc_groupe) # Limit to sets for 
# Filter the data : FOB data
obs_fob <- obs_clean %>%
  dplyr::filter(code_assoc_groupe == 1) 

# BIOLOGY DATA ----
samples_observe <- readr::read_csv("E:/git/reduce/data/samples_observe_FSCFOB.csv")
# Removes the year 2023 from the dataset and focuses solely on FOB schooltypes
samples_observe <- samples_observe %>%
  dplyr::filter(year != 2023,
                school_type == "FOB") %>%
  dplyr::mutate(month = format(as.Date(observation_date, format = "%Y-%m-%d"),
                               "%m"))
samples_observe$month = as.integer(samples_observe$month)

# Mauritania
samples_mauritania <- subset(samples_observe,
                             latitude >= 16 & latitude <= 21 & 
                               longitude >= -21 & longitude <= -16)

# Gabon
samples_gabon <- subset(samples_observe,
                        latitude <= 0 & latitude >= -4 & 
                          longitude >= 7 & longitude <= 10)

# Angola
samples_angola <- samples_observe %>%
  subset(latitude <= -7 & latitude >= -13 & 
           longitude >= 10 & longitude <= 14)

# Guinea
samples_guinea <- samples_observe %>%
  subset(latitude >= 0 & latitude <= 10 & 
           longitude >= -21 & longitude <= -10)

# Data clean from OBS FOB ----
# Select the bycatch species
data_clean_prep <- obs_fob %>%
  dplyr::select(c(obs_set_id,
           the_geom,
           heures_peche_standard,
           mois_de_debarquement,
           mois_de_peche,
           year,
           cwp11_geom,
           sunset,
           sunrise,
           obs_ts,
           # Billfishes
           num_BUM,
           num_SAI,
           # Other bony fishes
           num_ALN,
           num_ALM,
           num_CNT,
           num_CFW, 
           num_DIY,
           num_DOL,
           num_EHN, 
           num_GBA,
           num_JKY, 
           num_KYP,
           num_KYS,
           num_LOB,
           num_MOX,
           num_MRW, 
           num_REO,
           num_RRU,
           num_RUB,
           num_RZV,
           num_TRG,
           num_USE,
           num_WAH,
           num_YTL,
           # Rays
           num_PLS,
           num_RMM,
           num_RMT,
           # Sharks
           num_BSH,
           num_FAL,
           num_SPL,
           num_SPZ,
           num_SMA, 
           # Turtles
           num_LKV,
           num_TTL,
           # Tunas
           capture_skj_corrigee,
           capture_yft_corrigee,
           capture_bet_corrigee  ))

# Add longitude and latitude
data_clean_prep <- data_clean_prep %>%
  dplyr::mutate(longitude = sf::st_coordinates(the_geom)[,1],
                latitude = sf::st_coordinates(the_geom)[,2]) %>%
  sf::st_drop_geometry() # drop geometry only after extracting coordinates

# Replace NA values with 0 in numeric columns only
data_clean_prep <- data_clean_prep %>%
  dplyr::mutate(across(where(is.numeric), 
                       ~replace_na(., 0)))

# DATA CLEAN WITH SPECIES NAMES ----
data_clean <- data_clean_prep
# Change fao codes to group names and group species together
data_clean$Billfishes <- data_clean$num_BUM + data_clean$num_SAI
data_clean$Leatherjacket_filefishes <- data_clean$num_ALM + data_clean$num_ALN
data_clean$Whitespotted_filefishes <- data_clean$num_JKY
data_clean$Cottonmouth_jack <- data_clean$num_USE
data_clean$Rough_triggerfish <- data_clean$num_CNT
data_clean$Dolphinfishes <- data_clean$num_DOL + data_clean$num_CFW
data_clean$Great_barracuda <- data_clean$num_GBA
data_clean$Sea_chubs <- data_clean$num_KYP + data_clean$num_KYS
data_clean$Tripletail <- data_clean$num_LOB
data_clean$Sunfishes <- data_clean$num_MOX + data_clean$num_MRW + data_clean$num_RZV
data_clean$Sharksucker <-  data_clean$num_REO + data_clean$num_EHN
data_clean$Rainbow_runner <- data_clean$num_RRU
data_clean$Blue_runner <- data_clean$num_RUB
data_clean$Grey_triggerfish <- data_clean$num_TRG
data_clean$Wahoo <- data_clean$num_WAH
data_clean$Longfin_yellowtail <- data_clean$num_YTL
data_clean$Pelagic_stingray <- data_clean$num_PLS
data_clean$Devil_ray <- data_clean$num_RMM + data_clean$num_RMT
data_clean$Blue_shark <- data_clean$num_BSH
data_clean$Silky_shark <- data_clean$num_FAL
data_clean$Scalloped_hammerhead <-  data_clean$num_SPL
data_clean$Smooth_hammerhead <- data_clean$num_SPZ
data_clean$Shortfin_Mako <- data_clean$num_SMA
data_clean$Olive_ridley_turtle <- data_clean$num_LKV
data_clean$Loggerhead_turtle <- data_clean$num_TTL
data_clean$Major_tunas <- data_clean$capture_skj_corrigee + data_clean$capture_yft_corrigee + data_clean$capture_bet_corrigee

# Remove the num_ columns
data_clean <- data_clean %>%
  dplyr::select(-(tidyr::starts_with("num_")))

# Species select total
species_select_total <- data_clean_prep %>%
  dplyr::mutate(num_total = rowSums(select(., 
                                           tidyr::starts_with("num")))) 
# Works with time data
data_clean <- data_clean %>%
  dplyr::mutate(hrs_since_sunrise = difftime(obs_ts,
                                      sunrise, 
                                      units = "hours"),
         mins_since_sunrise = difftime(obs_ts,
                                       sunrise, 
                                       units = "mins"),
         hrs_since_sunrise = round(as.numeric(hrs_since_sunrise),
                                   digits = 0),
         quarter_since_sunrise = floor(as.numeric(mins_since_sunrise / 15)),
         obs_ts = as.POSIXct(obs_ts, 
                             format = "%Y-%m-%d %H:%M:%S", 
                             tz = "UTC"),
         sunrise = as.POSIXct(sunrise, 
                              format = "%Y-%m-%d %H:%M:%S", 
                              tz = "UTC"),
         sunset = as.POSIXct(sunset, 
                             format = "%Y-%m-%d %H:%M:%S",
                             tz = "UTC"),
         obs_dn = ifelse(obs_ts >= sunrise &
                           obs_ts <= sunset, 
                         "jour",
                         "nuit")) 
```

# *Dominant bycatch species and percentage caught on dFADs*

## *Table 1 - Table of the total observed number of bycatch individuals caught between 2013 and 2022*

```{r}
# Number of dFAD sets
n_dFAD_sets <- obs_clean %>%
  dplyr::filter(code_assoc_groupe == 1) %>%
  dplyr::distinct(obs_set_id) %>%
  nrow()
# Number of FSC sets
n_FSC_sets <- obs_clean %>%
  dplyr::filter(code_assoc_groupe == 2) %>%
  dplyr::distinct(obs_set_id) %>%
  nrow()
# FAO species order
fao_order <- c(# Billfishes
  "BUM", "SAI",
  # Other bony fishes
  "ALN", "ALM", "CNT", "CFW", "DIY", "DOL", "EHN", "GBA", "JKY",
  "KYP", "KYS", "LOB", "MOX", "MRW", "REO", "RRU", "RUB", "RZV",
  "TRG", "USE", "WAH", "YTL",
  # Rays
  "PLS", "RMM", "RMT",
  # Sharks
  "BSH", "FAL", "SPL", "SPZ", "SMA",
  # Turtles
  "LKV", "TTL")

obs_clean %>%
  dplyr::select(code_assoc_groupe, 
                tidyr::starts_with("num_")) %>%
  tidyr::pivot_longer(cols = starts_with("num_"),
                      names_to = "FAO",
                      names_prefix = "num_",
                      values_to = "count") %>%
  # Select only the species in 'fao_order'
  dplyr::mutate(FAO = factor(FAO, 
                             levels = fao_order)) %>%  
  dplyr::group_by(FAO) %>%
  dplyr::summarise(ALL = sum(count, 
                             na.rm = TRUE),
                   dFAD = sum(count[code_assoc_groupe == 1], 
                              na.rm = TRUE),
                   FSC = sum(count[code_assoc_groupe == 2],
                             na.rm = TRUE),
                   `no./dFAD set` = dFAD / n_dFAD_sets,
                   `no./FSC set` = FSC / n_FSC_sets,
                   `% caught on dFAD` = dFAD / (dFAD + FSC) * 100) %>%
  dplyr::arrange(FAO)  
```

### *Table S1 - Number of bycatch individuals per set, categorized by species group, between 2013 and 2022 on dFADs*

```{r, message=FALSE, warning=FALSE}
#| label: num-silky-per-year
#| echo: false
silky <- species_select_total %>%
  dplyr::group_by(year) %>%
  dplyr::summarize(n = sum(num_total, na.rm = TRUE),
                   n_fishes = sum(c(num_ALN,
                                    num_ALM,
                                    num_CNT,
                                    num_CFW, 
                                    num_DOL,
                                    num_EHN, 
                                    num_GBA,
                                    num_JKY, 
                                    num_KYP,
                                    num_KYS,
                                    num_LOB,
                                    num_MOX,
                                    num_MRW, 
                                    num_REO,
                                    num_RRU,
                                    num_RUB,
                                    num_RZV,
                                    num_TRG,
                                    num_USE,
                                    num_WAH,
                                    num_YTL), 
                                  na.rm = TRUE),
                   n_billfishes = sum(c(num_BUM,
                                        num_SAI),
                                      na.rm = TRUE),
                   n_rays = sum(c(num_PLS, 
                                  num_RMM,
                                  num_RMT),
                                na.rm = TRUE),
                   n_sharks = sum(c(num_BSH,
                                    num_FAL,
                                    num_SPL,
                                    num_SPZ,
                                    num_SMA), 
                                  na.rm = TRUE),
                   n_turtles = sum(c(num_LKV, 
                                     num_TTL), 
                                   na.rm = TRUE),
                   nsets = sum(!is.na(num_total) & num_total > 0),
                   nsets_tot = length(num_total) ) %>%
  dplyr::summarize(Prevalence = nsets / nsets_tot,
                   `Number per set` = n / nsets_tot,
                   `Number other bony fishes per set` = n_fishes / nsets_tot,
                   `Number billfishes per set` = n_billfishes / nsets_tot,
                   `Number rays per set` = n_rays / nsets_tot,
                   `Number sharks per set` = n_sharks / nsets_tot,
                   `Number turtles per set` = n_turtles / nsets_tot)

silky <- round(silky, 2)
silky
```

### *Table S2 - Summary statistics of species-specific catch, including the FAO species code, minimum, maximum, mean, and median number of individuals per set, as well as the minimum and maximum annual mean number of individuals per set. The colours are those of the IUCN protection levels used in table 1.*

```{r}
# List of columns corresponding to species
species_cols <- c("num_BUM", "num_SAI", "num_ALN", "num_ALM", "num_CNT", "num_CFW",
                  "num_DIY", "num_DOL", "num_EHN", "num_GBA", "num_JKY", "num_KYP",
                  "num_KYS", "num_LOB", "num_MOX", "num_MRW", "num_REO", "num_RRU",
                  "num_RUB", "num_RZV", "num_TRG", "num_USE", "num_WAH", "num_YTL",
                  "num_PLS", "num_RMM", "num_RMT", "num_BSH", "num_FAL", "num_SPL",
                  "num_SPZ", "num_SMA", "num_LKV", "num_TTL")

# Create a summary table for all species
summary_table <- species_cols %>%
  lapply(function(species) {
    # Compute yearly summary statistics for the current species
    silky_summary <- species_select_total %>%
      dplyr::group_by(year) %>%
      dplyr::summarize(n = sum(num_total, na.rm = TRUE),
                       n_species = sum(.data[[species]], na.rm = TRUE),
                       nsets = sum(!is.na(num_total) & num_total > 0),
                       nsets_tot = length(num_total),
                       .groups = "drop") %>%
      dplyr::mutate(Prevalence = nsets / nsets_tot,
                    `Number per set` = n / nsets_tot,
                    `Number species per set` = n_species / nsets_tot)
    # Return a tibble with overall summary statistics for the species
    tibble::tibble(FAO_Code = species,
                   min_ind_per_set = round(min(species_select_total[[species]],
                                               na.rm = TRUE), 2),
                   max_ind_per_set = round(max(species_select_total[[species]], 
                                               na.rm = TRUE), 2),
                   mean_ind_per_set = round(mean(species_select_total[[species]],
                                                 na.rm = TRUE), 2),
                   median_ind_per_set = round(median(species_select_total[[species]],
                                                     na.rm = TRUE), 2),
                   min_ind_per_year = round(min(silky_summary$`Number species per set`,
                                                na.rm = TRUE), 2),
                   max_ind_per_year = round(max(silky_summary$`Number species per set`,
                                                na.rm = TRUE), 2))}) %>%
  bind_rows()  # Combine all species rows into one final summary table

summary_table
```

### *Table S3 - Number of bycatch caught by species group between 2013 and 2022 on fishing aggregating devices*

```{r, message=FALSE, warning=FALSE}
#| echo: false
silky <- species_select_total %>%
  group_by(year) %>%
  summarize(n = sum(num_total, na.rm = TRUE),
            n_fishes = sum(c(num_ALN,
                             num_ALM,
                             num_CNT,
                             num_CFW,
                             num_DOL,
                             num_EHN, 
                             num_GBA,
                             num_JKY, 
                             num_KYP,
                             num_KYS,
                             num_LOB,
                             num_MOX,
                             num_MRW, 
                             num_REO,
                             num_RRU,
                             num_RUB,
                             num_RZV,
                             num_TRG,
                             num_USE,
                             num_WAH,
                             num_YTL), 
                           na.rm = TRUE),
            n_billfishes = sum(c(num_BUM,
                                 num_SAI),
                               na.rm = TRUE),
            n_rays = sum(c(num_PLS, 
                           num_RMM,
                           num_RMT),
                         na.rm = TRUE),
            n_sharks = sum(c(num_BSH,
                             num_FAL,
                             num_SPL,
                             num_SPZ,
                             num_SMA), 
                           na.rm = TRUE),
            n_turtles = sum(c(num_LKV, 
                              num_TTL), 
                            na.rm = TRUE),
            nsets = sum(!is.na(num_total) & num_total > 0),
            nsets_tot = length(num_total) )

silky <- round(silky, 2)
silky
```

# *Community structure*

## *Figure 1a - Dendrogram of bycatch species and IUCN conservation status based on NMDS and Jaccard dissimilarity (clusters were determined by a silhouette analysis)*

```{r}
# PRESENCE / ABSENCE DATA SETE ----
data_dendro <- data_clean[16:40]
data_dendro <- data_dendro[rowSums(data_dendro != 0) > 0, ]

# Convert the data to presence absence after fitering
presence_absence_matrix <- as.data.frame(lapply(data_dendro,
                                                function(x) as.integer(x != 0)))
species_occurrences <- colSums(presence_absence_matrix)

species_occurrences

# Outliers are removed from the dataset otherwise nmds and dissimilarity matrix will contains mistakes
cols_to_keep <- species_occurrences %>%
  names() %>%
  .[species_occurrences >= 84] #1% of occurence 

presence_absence_filtered <- presence_absence_matrix %>%
  select(all_of(cols_to_keep))

rows_to_keep <- rowSums(presence_absence_filtered != 0) > 0


#Filter both data frames using the same logical index
presence_absence_filtered <- presence_absence_filtered[rows_to_keep, ]

# Percentage deleted
total_occurrences <- sum(colSums(presence_absence_matrix))  # Total occurrences in the original data
kept_occurrences <- sum(colSums(presence_absence_filtered))  # Occurrences kept after filtering
percentage_deleted <- (total_occurrences - kept_occurrences) / total_occurrences * 100

# Output the results
cat("Total Occurrences:", 
    total_occurrences, "\n")
cat("Kept Occurrences:", 
    kept_occurrences, "\n")
cat("Percentage Deleted:",
    percentage_deleted, "%\n")

# Sites outliers
single_species_sites <- rowSums(presence_absence_filtered == 1) == 1
sites_with_one_species <- presence_absence_filtered[single_species_sites, ]

# Counting the number of such sites
num_sites_with_one_species <- nrow(sites_with_one_species)
cat("Number of sites with only one species present:", 
    num_sites_with_one_species, "\n")
# Focus on the bycatch community so we don't want fishing sets with 1 species
data_for_jaccard <- presence_absence_filtered[!single_species_sites, ]

# JACCARD ----
Jaccard_diss <- vegan::vegdist(data_for_jaccard,
                               method = "jaccard")
# NMDS ----
# Define the number of cores to use
num_cores <- min(parallel::detectCores() - 1, 4)  # Use up to 4 cores or one less than available
clus <- parallel::makeCluster(num_cores)

# Ensure that vegan functions are available on all cluster nodes
parallel::clusterEvalQ(clus, library(vegan))

# Run NMDS with parallel processing
nmds_results <- vegan::metaMDS(Jaccard_diss,
                                  distance = "jaccard",
                                  autotransform = FALSE,
                                  k = 3,
                                  trymax = 20,
                                  maxit = 999,
                                  parallel = clus) 
parallel::stopCluster(clus)

# CLUSTERING ----
# Hierarchical clustering on NMDS results
str(nmds_results)
stress_value <- nmds_results$stress
nmds_results$species <- wascores(nmds_results$points,
                                 data_for_jaccard,
                                 expand = TRUE) # you can either assign the results by sites or species
data <- nmds_results$species
nmds_scores <- as.data.frame(scores(nmds_results,
                                    display = "sites"))
# assign clusters to nmds results
diss <- dist(data)
spclusters <- hclust(diss)
clusters <- cutree(spclusters, k = 2)
clust.dend <- as.dendrogram(spclusters)
plot(spclusters)
```

## *Figure 1b - Co-occurrence Network to explore community structure and species associations.*

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Transpose the presence/absence table (species in columns and sites in rows)
t_data_for_jaccard <- t(data_for_jaccard)

# Compute the species co-occurrence matrix
co_occurrence_matrix <- t_data_for_jaccard %*% t(t_data_for_jaccard)
diag(co_occurrence_matrix) <- 0  # Set diagonals to zero to avoid self-links

# Calculate the sum of occurrences per species
species_sums <- rowSums(t_data_for_jaccard)

# # Normalize the co-occurrence matrix
# co_occurrence_normalized <- co_occurrence_matrix / outer(species_sums, 
#                                                          species_sums, 
#                                                          pmin)

# Méthode de Jaccard normalisée (recommandée)
union_matrix <- outer(species_sums, species_sums, "+") - co_occurrence_matrix
co_occurrence_normalized <- co_occurrence_matrix / union_matrix
diag(co_occurrence_normalized) <- 0

# Create the normalized network
species_network_normalized <- igraph::graph_from_adjacency_matrix(co_occurrence_normalized,
                                                                  mode = "undirected",
                                                                  weighted = TRUE, 
                                                                  diag = FALSE)

# Community detection on the normalized network
communities_normalized <- igraph::cluster_louvain(species_network_normalized)

# Louvain modularity
igraph::modularity(communities_normalized)

# Display the results
plot(species_network_normalized, 
     vertex.label = V(species_network_normalized)$name,
     vertex.color = communities_normalized$membership, 
     edge.width = E(species_network_normalized)$weight)

```

### *Figure S1 - Distribution maps of Group A (orange) and Group B (blue) species in the Atlantic, 2013-2022*

```{r}
# Now group by obs_set_id and retain longitude and latitude for each set
data_set_group <- data_clean_prep %>%
  dplyr::group_by(obs_set_id) %>%
  dplyr::summarize(dplyr::across(dplyr::starts_with("num_"), 
                          ~sum(.x, na.rm = TRUE)),
            longitude = first(longitude),  # retain longitude
            latitude = first(latitude)) %>%
  dplyr::mutate(across(tidyr::starts_with("num_"), 
                       ~ ifelse(.x > 1, 1, .x)))

# Define the two groups of species
group1_species <- c("num_DOL", "num_CFW",
                    "num_ALM", "num_ALN",
                    "num_TRG", 
                    "num_YTL", 
                    "num_RRU",
                    "num_USE",
                    "num_CNT", 
                    "num_RUB", 
                    "num_GBA", 
                    "num_WAH", 
                    "num_KYS", 
                    "num_KYP",
                    "num_LOB") 

group2_species <- c("num_SPZ",
                    "num_SMA",
                    "num_SPL",
                    "num_LKV", 
                    "num_FAL", 
                    "num_MOX", "num_MRW", "num_RZV",
                    "num_REO", "num_EHN",
                    "num_RMM", "num_RMT",
                    "num_BSH", 
                    "num_TTL",
                    "num_SAI", 
                    "num_BUM",
                    "num_PLS") 

# Add the percentages of species present in each group
data_set_group <- data_set_group %>%
  dplyr::rowwise() %>%
  dplyr::mutate(total_group1 = sum(dplyr::c_across(tidyr::all_of(group1_species)),
                                   na.rm = TRUE),
                total_group2 = sum(dplyr::c_across(tidyr::all_of(group2_species)),
                                   na.rm = TRUE),
                total_species = total_group1 + total_group2,
                perc_group1 = ifelse(total_species > 0, 
                                            total_group1 / total_species * 100, 0),
                perc_group2 = ifelse(total_species > 0,
                                            total_group2 / total_species * 100, 0))

data_set_group_filter <- data_set_group %>%
  dplyr::filter(perc_group1 == 100,
                total_group1 > 1)

plot1 <- ggplot2::ggplot() +
ggplot2::theme(legend.text = ggplot2::element_text(size = 10),
               panel.background = ggplot2::element_rect(fill = "white"),
               panel.border = ggplot2::element_rect(color = "black",
                                                    fill = NA,
                                                    size = 0.3))  +
  ggplot2::geom_sf(data = world_boundaries) +
  ggplot2::geom_point(data = data_set_group_filter, 
                      aes(x = longitude, y = latitude), 
                      color = "steelblue") +  
  ggplot2::scale_fill_viridis_c() +
  ggspatial::coord_sf(xlim = ocean_xlim,
                      ylim = ocean_ylim) +
  ggplot2::geom_hline(yintercept = ocean_yintercept,
                      linetype = "dashed",
                      color = "darkgrey",
                      linewidth = 0.2) +
  ggplot2::labs(fill = "Number ind.",
                caption = "",
                x = "",
                y = "") +
  ggplot2::geom_vline(xintercept = ocean_xintercept,
                      linetype = "dashed",
                      color = "darkgrey",
                      linewidth = 0.2)

data_set_group_filter <- data_set_group %>%
  dplyr::filter(perc_group2 == 100,
         total_group2 > 1)


plot2 <- ggplot2::ggplot() +
ggplot2::theme(legend.text = ggplot2::element_text(size = 10),
               panel.background = ggplot2::element_rect(fill = "white"),
               panel.border = ggplot2::element_rect(color = "black",
                                                    fill = NA,
                                                    size = 0.3))  +
  ggplot2::geom_sf(data = world_boundaries) +
  ggplot2::geom_point(data = data_set_group_filter, 
                      aes(x = longitude, y = latitude), 
                      color = "orange") +  
  ggplot2::scale_fill_viridis_c() +
  ggspatial::coord_sf(xlim = ocean_xlim,
                      ylim = ocean_ylim) +
  ggplot2::geom_hline(yintercept = ocean_yintercept,
                      linetype = "dashed",
                      color = "darkgrey",
                      linewidth = 0.2) +
  ggplot2::labs(fill = "Number ind.",
                caption = "",
                x = "",
                y = "") +
  ggplot2::geom_vline(xintercept = ocean_xintercept,
                      linetype = "dashed",
                      color = "darkgrey",
                      linewidth = 0.2)

plot1 + plot2


```

### *Figure S2 - Number of species caught per hour (in hours since sunrise) of vulnerable species group (orange) and silky billfishes group (blue) species in the Atlantic by sets and by tunas*

```{r}
# Data clean from OBS FOB ----
# DATA CLEAN WITH SPECIES NAMES ----
data_clean_dn <- data_clean %>%
  # Creation of the three groups
  dplyr::mutate("group_vulnerable_species" = Loggerhead_turtle + Sharksucker + Smooth_hammerhead + Pelagic_stingray +
           Devil_ray + Scalloped_hammerhead + Blue_shark + Sunfishes + Olive_ridley_turtle,
         "group_silky_billfishes" = Billfishes + Silky_shark) %>%
  # Group by hours since sunrise
  dplyr::group_by(hrs_since_sunrise) %>%
  dplyr::summarise("group_vulnerable_species" = sum(group_vulnerable_species, 
                                                    na.rm = TRUE),
            "group_silky_billfishes" = sum(group_silky_billfishes, 
                                           na.rm = TRUE),
            "nb_sets" = dplyr::n(),
            "nb_tons_tunas" = sum(Major_tunas, 
                                  na.rm = TRUE),
            "total" = group_vulnerable_species + group_silky_billfishes,
            .groups = "drop") %>%
  # Calculating the percentage
  dplyr::mutate("VS_by_sets" = round((group_vulnerable_species / nb_sets), 2),
         "VS_by_tunas" = round((group_vulnerable_species / nb_tons_tunas), 2),
         "silky_by_sets" = round((group_silky_billfishes / nb_sets), 2),
         "silky_by_tunas" = round((group_silky_billfishes / nb_tons_tunas), 2))

# Data normality test - Shapiro (H0 = normal data)
shapiro.test(data_clean_dn$VS_by_sets) # Normal data
shapiro.test(data_clean_dn$VS_by_tunas) # Normal data
shapiro.test(data_clean_dn$silky_by_sets)# Normal data
shapiro.test(data_clean_dn$silky_by_tunas) # Normal data

# Variance homogeneity (H0 = Variance not homogeneous)
# car::leveneTest((Loggerhead_turtle + Sharksucker + Smooth_hammerhead + Pelagic_stingray +
#                    Devil_ray + Scalloped_hammerhead + Blue_shark + Sunfishes + Olive_ridley_turtle) ~ as.factor(hrs_since_sunrise), data = data_clean) # homogeneous
# car::leveneTest((Billfishes + Silky_shark) ~ as.factor(hrs_since_sunrise), data = data_clean) #homogeneous
# 
# car::leveneTest(((Loggerhead_turtle + Sharksucker + Smooth_hammerhead + Pelagic_stingray +
#                    Devil_ray + Scalloped_hammerhead + Blue_shark + Sunfishes + Olive_ridley_turtle) / Major_tunas) ~ as.factor(hrs_since_sunrise ), data = data_clean) #not homogeneous
# car::leveneTest((Billfishes + Silky_shark) / Major_tunas ~ as.factor(hrs_since_sunrise ), data = data_clean) #homogeneous

# anova
# anova_vs_sets <- aov((Loggerhead_turtle + Sharksucker + Smooth_hammerhead + Pelagic_stingray +
#                          Devil_ray + Scalloped_hammerhead + Blue_shark + Sunfishes + Olive_ridley_turtle) ~ as.factor(hrs_since_sunrise), data = data_clean)
# summary(anova_vs_tunas)
# TukeyHSD(anova_vs_tunas)
# Non-significant anova
# 
# anova_vs_tunas <- aov((Loggerhead_turtle + Sharksucker + Smooth_hammerhead + Pelagic_stingray +
#                          Devil_ray + Scalloped_hammerhead + Blue_shark + Sunfishes + Olive_ridley_turtle / Major_tunas) ~ as.factor(hrs_since_sunrise), data = data_clean)
# summary(anova_vs_tunas)
# TukeyHSD(anova_vs_tunas)
# # Non-significant anova
# 
# anova_silky_tunas <- aov((Billfishes + Silky_shark/ Major_tunas) ~ as.factor(hrs_since_sunrise), data = data_clean)
# summary(anova_vs_tunas)
# TukeyHSD(anova_vs_tunas)
# # Non-significant anova
# 
# anova_silky_sets <- aov((Billfishes + Silky_shark) ~ as.factor(hrs_since_sunrise ), data = data_clean)
# summary(anova_silky_tunas)
# TukeyHSD(anova_silky_tunas)
# # Non-significant anova

# Filter to exclude time 12
data_clean_dn <- data_clean_dn %>%
  dplyr::filter(hrs_since_sunrise != 12)

# Colour palette
colors <- c("Vulnerable species group" = "orange", 
            "Silky-billfishes group" = "steelblue")

# VS_by_sets
plot1 <- ggplot2::ggplot(data_clean_dn, aes(x = as.factor(hrs_since_sunrise),
                                            y = VS_by_sets,
                                            fill = "Vulnerable species group")) +
  ggplot2::geom_col() +
  ggplot2::geom_text(ggplot2::aes(label = nb_sets), 
                     vjust = -0.5, 
                     size = 3.5) +
  ggplot2::scale_fill_manual(values = colors, name = "") +
  ggplot2::labs(x = "", 
                y = "Nb. of ind. caught by sets", 
                title = "") +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::theme(legend.position = "none") +
  ggplot2::coord_cartesian(ylim = c(0, 2.7))

# silky_by_sets
plot2 <- ggplot2::ggplot(data_clean_dn, 
                         ggplot2::aes(x = as.factor(hrs_since_sunrise), 
                                      y = silky_by_sets, 
                                      fill = "Silky-billfishes group")) +
  ggplot2::geom_col() +
  ggplot2::geom_text(aes(label = nb_sets), 
                     vjust = -0.5, 
                     size = 3.5) +
  ggplot2::scale_fill_manual(values = colors, name = "") +
  ggplot2::labs(x = "", 
                y = "", 
                title = "") +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::theme(legend.position = "none")+
  ggplot2::coord_cartesian(ylim = c(0, 2.7))

# VS_by_tunas
plot3 <- ggplot2::ggplot(data_clean_dn, 
                         ggplot2::aes(x = as.factor(hrs_since_sunrise), 
                                      y = VS_by_tunas, 
                                      fill = "Vulnerable species group")) +
  ggplot2::geom_col() +
  ggplot2::geom_text(ggplot2::aes(label = nb_sets), 
                     vjust = -0.5, 
                     size = 3.5) +
  ggplot2::scale_fill_manual(values = colors, 
                             name = "") +
  ggplot2::labs(x = "Hours since sunrise", 
                y = "Nb. of ind. caught by tons of major tunas",
                title = "") +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::theme(legend.position = "none")+
  ggplot2::coord_cartesian(ylim = c(0, 0.15))

# silky_by_tunas
plot4 <- ggplot2::ggplot(data_clean_dn, 
                         ggplot2::aes(x = as.factor(hrs_since_sunrise),
                                      y = silky_by_tunas, 
                                      fill = "Silky-billfishes group")) +
  ggplot2::geom_col() +
  ggplot2::geom_text(ggplot2::aes(label = nb_sets),
                     vjust = -0.5, 
                     size = 3.5) +
  ggplot2::scale_fill_manual(values = colors, 
                             name = "") +
  ggplot2::labs(x = "Hours since sunrise",
                y = "", 
                title = "") +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::theme(legend.position = "none")+
  ggplot2::coord_cartesian(ylim = c(0, 0.15))

# Assembling graphs with a single common legend
legend <- cowplot::get_legend(plot1 + 
                                ggplot2::theme(legend.position = "bottom"))

final_plot <- (plot1 | plot2) / (plot3 | plot4) + 
  patchwork::plot_layout(guides = "collect") & 
  ggplot2::theme(legend.position = "bottom")

cowplot::plot_grid(final_plot, 
                   legend, ncol = 1, 
                   rel_heights = c(1, 0.1))
```

# *Hotspot and distribution of bycatch species*

## *Figure 2a - Bycatch hotspots of vulnerable species in the Eastern Atlantic with top species per hotspot*

### *Map*

```{r}
bs <- data_clean_prep %>%
  dplyr::select(c(the_geom,
           mois_de_debarquement,
           num_BUM,
           num_MOX,
           num_RMM,
           num_RMT,
           num_BSH,
           num_FAL,
           num_SPL,
           num_SPZ,
           num_SMA,
           num_LKV,
           num_TTL)) %>%
  dplyr::mutate(num_TOTAL = rowSums(dplyr::select(., 
                                                  tidyr::starts_with("num_")), 
                                    na.rm = TRUE)) %>%
  dplyr::filter(num_TOTAL > 20, #Minimum 20 individuals per set
         num_TOTAL != 190) %>% # Exclude the 190 ind. outlier for a better visibility
  dplyr::arrange(num_TOTAL)

bs = sf::st_as_sf(bs)

bs <- bs %>%
  dplyr::mutate(longitude = sf::st_coordinates(the_geom)[,1],
         latitude = sf::st_coordinates(the_geom)[,2]) %>%
  sf::st_drop_geometry()


## Now plot the results
(map_year <- ggplot2::ggplot() +
    ggplot2::theme(legend.text = ggplot2::element_text(size = 10),
                   panel.background = ggplot2::element_rect(fill = "white"),
                   panel.border = ggplot2::element_rect(color = "black",
                                                        fill = NA,
                                                        size = 0.3),
                   legend.position = c(0.13, 0.12)) +  # Moving the legend to the bottom left
    ggplot2::geom_sf(data = world_boundaries) +
    ggplot2::geom_point(data = bs, 
                        aes(x = longitude, 
                            y = latitude, 
                            fill = num_TOTAL),
                        size = 3, 
                        shape = 21) +  
    ggplot2::scale_fill_viridis_c() +
    ggspatial::coord_sf(xlim = ocean_xlim,
                        ylim = ocean_ylim) +
    ggplot2::geom_hline(yintercept = ocean_yintercept,
                        linetype = "dashed",
                        color = "darkgrey",
                        linewidth = 0.2) +
    ggplot2::labs(fill = "Number ind.",
                  caption = "",
                  x = "",
                  y = "") +
    ggplot2::geom_vline(xintercept = ocean_xintercept,
                        linetype = "dashed",
                        color = "darkgrey",
                        linewidth = 0.2))
```

### *Species composition*

```{r}
data_clean_pie <- data_clean %>%
  dplyr::select(c(latitude,
           longitude,
           Billfishes,
           Sunfishes,
           Devil_ray,
           Blue_shark,
           Silky_shark,
           Scalloped_hammerhead,
           Smooth_hammerhead, 
           Shortfin_Mako,
           Olive_ridley_turtle,
           Loggerhead_turtle))

# Separation of data sets by hotspot

# MAURITANIE ----
hotspot_mauritania <- subset(data_clean_pie,
                             latitude >= 16 & latitude <= 21 & 
                               longitude >= -21 & longitude <= -16)
pie_chart_species_composition(hotspot_mauritania)

# GABON ----
hotspot_gabon <- subset(data_clean_pie,
                        latitude <= 0 & latitude >= -4 & 
                          longitude >= 7 & longitude <= 10)

pie_chart_species_composition(hotspot_gabon)

# ANGOLA ----
hotspot_angola <- subset(data_clean_pie,
                         latitude <= -7 & latitude >= -13 & 
                           longitude >= 10 & longitude <= 14)
pie_chart_species_composition(hotspot_angola)



# GUINEA ----
hotspot_guinea <- subset(data_clean_pie,
                         latitude >= 0 & latitude <= 10 & 
                           longitude >= -21 & longitude <= -10)
pie_chart_species_composition(hotspot_guinea)

# Atlantique ----
pie_chart_species_composition(data_clean_pie)
```

## *Figure 2b - Breakdown by sex and life-stage of the three top shark species caught in each hotspot and the Atlantic overall*

```{r}
# Size juvenile
#size_juv <- 225 # Silky shark
#size_juv <- 230 # Scalloped hammerhead
#size_juv <- 265 # Smooth hammerhead
size_juv <- 200 # Rays
#size_juv <- 221 # Blue shark


samples_observe_bio <- samples_observe[!(
  # Exclude data from the first area (latitude between 16 and 21, longitude between -21 and -16)
  (samples_observe$latitude >= 16 & samples_observe$latitude <= 21 & 
     samples_observe$longitude >= -21 & samples_observe$longitude <= -16) |
  
  # Exclude data from the second area (latitude between 0 and -4, longitude between 7 and 10)
  (samples_observe$latitude <= 0 & samples_observe$latitude >= -4 & 
     samples_observe$longitude >= 7 & samples_observe$longitude <= 10) |
  
  # Exclude data from the third area (latitude between -7 and -13, longitude between 10 and 14)
  (samples_observe$latitude <= -7 & samples_observe$latitude >= -13 & 
     samples_observe$longitude >= 10 & samples_observe$longitude <= 14) |
  
  # Exclude data from the area between latitude 0-10 and longitude -21 to -10
  (samples_observe$latitude >= 0 & samples_observe$latitude <= 10 & 
     samples_observe$longitude >= -21 & samples_observe$longitude <= -10)
), ]

bio_observe <- samples_observe_bio %>%
  dplyr::filter(fao_code %in% c("SPL")) %>%
  dplyr::select(fao_code,
         length, 
         size_type, 
         weight,
         count,
         sex, 
         year, 
         set_id)

bio_mauritania <- samples_mauritania %>%
  dplyr::filter(fao_code %in% c("SPL")) %>%
  dplyr::select(fao_code, 
         length, 
         size_type,
         weight,
         count,
         sex, 
         year, 
         set_id)

bio_gabon <- samples_gabon %>%
  dplyr::filter(fao_code %in% c("SPL")) %>%
  dplyr::select(fao_code, 
         length, 
         size_type, 
         weight, 
         count, 
         sex, 
         year,
         set_id)

bio_angola <- samples_angola %>%
  dplyr::filter(fao_code %in% c("SPL")) %>%
  dplyr::select(fao_code, 
         length, 
         size_type, 
         weight, 
         count, 
         sex, 
         year, 
         set_id)

bio_guinea <- samples_guinea %>%
  dplyr::filter(fao_code %in% c("SPL")) %>%
  dplyr::select(fao_code,
         length, 
         size_type, 
         weight, 
         count, 
         sex, 
         year, 
         set_id)

# CHI2 ----
# Apply the summarize_data function to each dataset
result_observe <- summarize_data(bio_observe, 
                                         size_juv)
result_mauritania <- summarize_data(bio_mauritania,
                                            size_juv)
result_gabon <- summarize_data(bio_gabon,
                                       size_juv)
result_angola <- summarize_data(bio_angola,
                                        size_juv)
result_guinea <- summarize_data(bio_guinea,
                                        size_juv)

# Add hotspot names
result_observe$Hotspot <- "Atlantic"
result_mauritania$Hotspot <- "Mauritania"
result_gabon$Hotspot <- "Gabon"
result_angola$Hotspot <- "Angola"
result_guinea$Hotspot <- "Guinea"

# Combine the results
final_result_SPL_FOB <- rbind(result_mauritania, 
                              result_gabon, 
                              result_angola,
                              result_guinea,
                              result_observe)
final_result_SPL_FOB

# Create the contingency table
tableau_contingence <- final_result_SPL_FOB %>%
  tidyr::pivot_longer(-Hotspot, 
               names_to = "Category",
               values_to = "Number") %>%
  tidyr::pivot_wider(names_from = Hotspot, 
              values_from = Number)

# Convert to matrix for Chi-squared test
matrice_contingence <- tableau_contingence %>%
  tibble::column_to_rownames(var = "Category") %>% # Set first column as row names
  as.matrix()

# Compute expected values
expected <- chisq.test(matrice_contingence)$expected
expected # Should be greater than 5 in 80% of the cells, otherwise use Fisher’s test

# Perform Chi-squared test
test_chi2 <- chisq.test(matrice_contingence)
test_chi2

# Extract standardized residuals
residus <- test_chi2$stdres
df_residus <- as.data.frame(as.table(residus))

# Heatmap of standardized residuals
ggplot2::ggplot(df_residus, 
       ggplot2::aes(Var2, 
           Var1, 
           fill = Freq)) +
  ggplot2::geom_tile() +  # Heatmap
  ggplot2::geom_text(aes(label = round(Freq, 2)), 
                     color = "black") +  # Add rounded values
  ggplot2::scale_fill_gradient2(low = "#023e8a", 
                                mid = "white",
                                high = "#d90429") +
  ggplot2::labs(title = "Standardized residuals of Chi² test - SPL",
       x = "Hotspot", y = "Category") +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position = "none")  # Remove legend

# CHART PIE ----

# Apply the function to each dataset
percentage_observe <- summarize_percentage(bio_observe,
                                                   size_juv)
percentage_mauritania <- summarize_percentage(bio_mauritania,
                                                      size_juv)
percentage_gabon <- summarize_percentage(bio_gabon, 
                                                 size_juv)
percentage_angola <- summarize_percentage(bio_angola, 
                                                  size_juv)
percentage_guinea <- summarize_percentage(bio_guinea, 
                                                  size_juv)

# Plot without legend
plot_observe <- pie_chart_percentage_bio(percentage_observe,
                                         "Atlantic", 
                                         legend = FALSE)
plot_mauritania <- pie_chart_percentage_bio(percentage_mauritania,
                                            "Mauritania", 
                                            legend = FALSE)
plot_gabon <- pie_chart_percentage_bio(percentage_gabon, 
                                       "Gabon", 
                                       legend = FALSE)
plot_angola <- pie_chart_percentage_bio(percentage_angola, 
                                        "Angola", 
                                        legend = FALSE)
plot_guinea <- pie_chart_percentage_bio(percentage_guinea, 
                                        "Guinea", 
                                        legend = FALSE)

# Create a plot to extract a single legend
legend_plot <- ggplot2::ggplot(percentage_observe %>%
                        dplyr::select(-total_ind) %>%
                        tidyr::pivot_longer(cols = tidyr::everything(),
                                     names_to = "Category", 
                                     values_to = "Percentage"),
                      ggplot2::aes(x = "", 
                          y = Percentage,
                          fill = Category)) +
  ggplot2::geom_bar(stat = "identity", 
           width = 1) +
  ggplot2::scale_fill_manual(values = c(perc_male_adult = "#005f73",
                               perc_male_juvenile = "#0a9396",
                               perc_female_adult = "#B56B3C",
                               perc_female_juvenile = "#e9d8a6")) +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position = "right",
        axis.text = ggplot2::element_blank(),
        axis.ticks = ggplot2::element_blank(),
        panel.grid = ggplot2::element_blank()) +
  ggplot2::labs(fill = "Category")

# Extract the single legend
get_legend <- function(my_plot) {
  tmp <- ggplot2::ggplot_gtable(ggplot2::ggplot_build(my_plot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- patchwork::wrap_elements(get_legend(legend_plot))

# Arrange the plots in 2x2 (or more) layout with legend on the right
layout <- (plot_observe + 
             plot_mauritania + 
             plot_gabon + 
             plot_angola + 
             plot_guinea) + 
  patchwork::plot_layout(guides = "collect") & 
  ggplot2::theme(legend.position = "right")

# Display the final plot
layout

# Tracking table for % of juveniles, males, and females by species ----

# Add hotspot names
percentage_observe$Hotspot <- "Atlantic"
percentage_mauritania$Hotspot <- "Mauritania"
percentage_gabon$Hotspot <- "Gabon"
percentage_angola$Hotspot <- "Angola"
percentage_guinea$Hotspot <- "Guinea"

# Combine the results
final_result_SPL_FOB <- rbind(percentage_mauritania,
                              percentage_gabon, 
                              percentage_angola, 
                              percentage_guinea,
                              percentage_observe)

# Display the final table
final_result_SPL_FOB

```

### *Figure S3: Count of the number of sets per month and per hotspot region*

```{r}
bs <- obs_fob %>%
  dplyr::select(c(the_geom,
                  obs_set_id,
                  mois_de_peche,
                  nombre_de_calees_pos)) 

bs <- bs %>%
  dplyr::mutate(longitude = sf::st_coordinates(the_geom)[,1],
         latitude = sf::st_coordinates(the_geom)[,2]) %>%
  sf::st_drop_geometry()

bs <- bs %>%
  dplyr::filter(latitude >= 16 & latitude <= 21 & longitude >= -21 & longitude <= -16, # Mauritanie
         #latitude <= 0 & latitude >= -4 & longitude >= 7 & longitude <= 10, #Gabon
         #latitude <= -7 & latitude >= -13 & longitude >= 10 & longitude <= 14, # Angola
         #latitude >= 0 & latitude <= 10 & longitude >= -21 & longitude <= -10 # Guinée
         )

data_set_group <- bs %>%
  dplyr::group_by(obs_set_id, mois_de_peche) %>%
  dplyr::summarize(nombre_de_calees = sum(nombre_de_calees_pos, 
                                   na.rm = TRUE),  # sum of set by obs_set_id
            longitude = first(longitude),  # keeps the first longitude value
            latitude = first(latitude))

data_set_group$mois_de_peche = factor(data_set_group$mois_de_peche,levels=1:12,ordered=TRUE)

ggplot2::ggplot(data=data_set_group,
                         ggplot2::aes(x = mois_de_peche)) +
  ggplot2::stat_count() +
  ggplot2::scale_x_discrete(drop=FALSE) +
  ggplot2::labs(x = "Month", y = "count of total set") +
  ggplot2::theme_bw() +
  ggplot2::ylim(0, 700) +
  ggplot2::geom_text(stat = 'count', 
                     ggplot2::aes(label = ..count..), 
                     vjust = -0.5)

```

### *Table S6 - Number of species per sets observed in the biodiversity hotspot of Mauritania.*

```{r}
resultat <- data_clean_prep %>%
  dplyr::filter(latitude >= 16 & latitude <= 21 & longitude >= -21 & longitude <= -16) %>%
  dplyr::mutate(num_RM = num_RMM + num_RMT) %>%  # Replaces num_RMM and num_RMT with num_RM
  dplyr::select(num_BUM,
                num_MOX,
                num_TRG,
                num_RM, 
                num_BSH,
                num_FAL,
                num_SPL,
                num_SPZ,
                num_SMA, 
                num_LKV,
                num_TTL,
                mois_de_peche) %>%
  dplyr::group_by(mois_de_peche) %>%
  dplyr::summarise(dplyr::across(tidyr::starts_with("num_"),
                                 sum,
                                 na.rm = TRUE)) 

nb_lignes_par_mois <- data_clean_prep %>%
  dplyr::filter(latitude >= 16 & latitude <= 21 & longitude >= -21 & longitude <= -16) %>%
  dplyr::group_by(mois_de_peche) %>%
  dplyr::summarise(Nombre_lignes = dplyr::n())

resultat_nb <- resultat %>%
  dplyr::left_join(nb_lignes_par_mois, by = "mois_de_peche")

resultat_final <- resultat_nb %>%
  dplyr::mutate(round(dplyr::across(dplyr::starts_with("num_"),
                                    ~ .x / Nombre_lignes), 2))

resultat_final
```

## *Table S7 -Number of species per sets observed in the biodiversity hotspot of Gabon*

```{r}
resultat <- data_clean_prep %>%
  dplyr::filter(latitude <= 0 & latitude >= -4 & longitude >= 7 & longitude <= 10) %>%
  dplyr::mutate(num_RM = num_RMM + num_RMT) %>%  # Replaces num_RMM and num_RMT with num_RM
  dplyr::select(num_BUM,
         num_MOX,
         num_TRG,
         num_RM,  
         num_BSH,
         num_FAL,
         num_SPL,
         num_SPZ,
         num_SMA, 
         num_LKV,
         num_TTL,
         mois_de_peche) %>%
  dplyr::group_by(mois_de_peche) %>%
  dplyr::summarise(dplyr::across(tidyr::starts_with("num_"), 
                                 sum, 
                                 na.rm = TRUE)) 



nb_lignes_par_mois <- data_clean_prep %>%
  dplyr::filter(latitude <= 0 & latitude >= -4 & longitude >= 7 & longitude <= 10) %>%
  dplyr::group_by(mois_de_peche) %>%
  dplyr::summarise(Nombre_lignes = dplyr::n())

resultat_nb <- resultat %>%
  dplyr::left_join(nb_lignes_par_mois, by = "mois_de_peche")

resultat_final <- resultat_nb %>%
  dplyr::mutate(round(dplyr::across(dplyr::starts_with("num_"), 
                                    ~ .x / Nombre_lignes), 2))


resultat_final

```

### *Table S8 - Number of species per sets observed in the biodiversity hotspot of Angola*

```{r}
resultat <- data_clean_prep %>%
  dplyr::filter(latitude <= -7 & latitude >= -13 & longitude >= 10 & longitude <= 14) %>%
  dplyr::mutate(num_RM = num_RMM + num_RMT) %>%  # Replaces num_RMM and num_RMT with num_RM
  dplyr::select(num_BUM,
         num_MOX,
         num_TRG,
         num_RM,  
         num_BSH,
         num_FAL,
         num_SPL,
         num_SPZ,
         num_SMA, 
         num_LKV,
         num_TTL,
         mois_de_peche) %>%
  dplyr::group_by(mois_de_peche) %>%
  dplyr::summarise(dplyr::across(dplyr::starts_with("num_"), 
                                 sum, 
                                 na.rm = TRUE)) 



nb_lignes_par_mois <- data_clean_prep %>%
  dplyr::filter(latitude <= -7 & latitude >= -13 & longitude >= 10 & longitude <= 14) %>%
  dplyr::group_by(mois_de_peche) %>%
  dplyr::summarise(Nombre_lignes = dplyr::n())

resultat_nb <- resultat %>%
  dplyr::left_join(nb_lignes_par_mois, by = "mois_de_peche")

resultat_final <- resultat_nb %>%
  dplyr::mutate(round(dplyr::across(dplyr::starts_with("num_"), ~ .x / Nombre_lignes), 2))


resultat_final
```

## *Figure 3 - Spatial and Temporal Distribution of Smooth Hammerhead (SPZ) in Mauritania.*

```{r}
# Mauritania
generate_fishing_maps_and_charts(obs_fob,
                                 world_boundaries,
                                 hotspot = "mauritania",
                                 species = "SPZ")
```

### *Figure S4 - Spatial and Temporal Distribution of Scalloped Hammerhead (SPL) in Mauritania.*

```{r}
# Gabon
generate_fishing_maps_and_charts(obs_fob,
                                 world_boundaries,
                                 hotspot = "gabon",
                                 species = "SPZ")

```

### *Figure S5 - Spatial and Temporal Distribution of Silky shark (FAL) in Gabon.*

```{r}
# Angola
generate_fishing_maps_and_charts(obs_fob,
                                 world_boundaries,
                                 hotspot = "angola",
                                 species = "FAL")
```

### *Figure S6 - Sampling effort*

```{r}
y = data_clean
y <- st_as_sf(data_clean, coords = c("longitude",
                                     "latitude"), 
              crs = 4326)  # CRS = WGS84


z <- st_make_grid(y, cellsize = 0.5)

z = st_make_grid(y,cellsize=0.5)
zz = st_intersects(z,y)
zzz = apply(zz,1,sum)
df = data.frame(geom=z,count=zzz)
df = st_as_sf(df)

eez <- readRDS("E:/git/reduce/inst/eez.rds")

worldmap = rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")

ggplot2::ggplot() +
  ggplot2::geom_sf(data = df, 
                   ggplot2::aes(fill = count), 
                   color = NA) +
  ggplot2::scale_fill_gradientn(colours = grDevices::colorRampPalette(c("white", "red"))(100),
                                name = "Sampling effort",  
                                breaks = scales::pretty_breaks(n = 5), # number of ticks on the legend
                                labels = scales::comma_format()) +
  ggplot2::geom_sf(data = eez, color = "black", fill = NA) +
  ggplot2::geom_sf(data = worldmap, fill = "grey90", color = "grey60", size = 0.1) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(panel.grid.major = ggplot2::element_line(color = "grey80", size = 0.1),
                 legend.background = ggplot2::element_rect(fill = "gray95", color = NA),
                 legend.key = ggplot2::element_rect(fill = "gray95", color = NA),
                 panel.background = ggplot2::element_rect(fill = "white", color = NA),
                 legend.position = "right",
                 legend.title = ggplot2::element_text(size = 12, face = "bold", color = "black"),
                 legend.text = ggplot2::element_text(size = 10, color = "black"),
                 legend.key.height = ggplot2::unit(2, "cm"),  # height of the colour bar
                 legend.key.width = ggplot2::unit(0.5, "cm")) +
  ggplot2::guides(fill = ggplot2::guide_colorbar(barwidth = 0.5, 
                                                 barheight = 8,
                                                 ticks = TRUE,
                                                 ticks.colour = "black",
                                                 frame.colour = "black",
                                                 frame.linewidth = 0.5)) +
  ggplot2::coord_sf(xlim = sf::st_bbox(df)[c("xmin", "xmax")],
                    ylim = sf::st_bbox(df)[c("ymin", "ymax")],
                    expand = FALSE) +
  ggplot2::geom_vline(xintercept = ocean_xintercept,
                      linetype = "dashed",
                      color = "darkgrey",
                      linewidth = 0.2) +
  ggplot2::geom_hline(yintercept = ocean_yintercept,
                      linetype = "dashed",
                      color = "darkgrey",
                      linewidth = 0.2) +
  ggplot2::theme(panel.border = ggplot2::element_rect(color = "black",
                                                      fill = NA, 
                                                      size = 0.3))

```
